---
title: "Lecture9.18"
output: github_document
---
```{r results='hide', message=FALSE}
library(readr)
library(tidyverse)
```
# Data Manipulation
First, import the data set & clean the variable names.
```{r collapse=TRUE, results='hide', message=FALSE}
litters_df = read_csv("../../data_import_examples/FAS_litters.csv", na = c("NA", ".", ""))
litters_df = janitor::clean_names(litters_df)
names(litters_df)
```
## `select` variables from dataframe
We apply `select` on **columns**.

* select specific columns
```{r results='hide'}
select(litters_df, group, litter_number, gd0_weight)
```
* select a range of columns
```{r results='hide'}
select(litters_df, group:gd_of_birth)
```
* select by remove certain columns
```{r results='hide'}
select(litters_df, -group, -litter_number, -gd0_weight)
select(litters_df, -(group:gd_of_birth))
```
* select by prefix
```{r results='hide'}
select(litters_df, group, starts_with("gd"))
```
* select by suffix
```{r results='hide'}
select(litters_df, group, contains("gd"))
```
* rearrange by selecting
```{r results='hide'}
select(litters_df, starts_with("gd"), group)
select(litters_df, litter_number, everything()) 
```
put `litternumber` piror to else, same to 
```{r results='hide'}
relocate(litters_df, litter_number)
```
* rename by selecting 
```{r results='hide'}
#rename group variable
select(litters_df, GROUP = group, everything())
rename(litters_df, GROUP = group) # same function
```
## `filter` 
We use `filter` to filter out **rows**.

* `filter` will turn rows that return **TRUE** to all conditions
```{r results='hide'}
filter(litters_df, gd_of_birth == 20, pups_born_alive >= 6, pups_dead_birth <= 0)
```
* filter for non-matches
```{r results='hide'}
filter(litters_df, pups_born_alive !=6)
```
* match to a set {}
```{r results='hide'}
# return group that in Condition7 and Condition8
filter(litters_df, group %in% c("Con7", "Con8"))
```
* `filter` out missing rows with `drop_na()`
```{r results='hide'}
drop_na(litters_df) #for the entire data set
drop_na(litters_df, gd0_weight) #for specific rows
```
# `mutate`
